---
title: "R packages authored by ACEMS Members"
author: "Rob J Hyndman"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  error = FALSE
)
library(tidyverse)
library(stringi)
library(glue)
library(cranlogs)
```

```{r acemspackages, include=FALSE}
# Download meta data on all CRAN packages
# and extract maintainer names
# download.file("http://cran.r-project.org/web/packages/packages.rds", "packages.rds");
allcran <- as.tibble(readRDS("packages.rds"), row.names = NA) %>%
  mutate(
    Family = str_extract(Maintainer, "[a-zA-Z]* <"),
    Family = str_extract(Family, "[a-zA-Z]*"),
    Given = str_extract(Maintainer, "[A-Za-z]*"),
    GivenInitials = str_sub(Given, 1, 2)
  )

# Download list of all ACEMS members
# acems <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1g1HP35bvlaOdDmx5zvDkI2e1diHOePkzgQPaEqmGnLk/edit?usp=sharing") %>%
acems <- read_csv("ACEMS_Researcher_ID_numbers - ACEMS_Researcher_ID_numbers.csv") %>%
  rename(institution = "Organisational Unit") %>%
  mutate(GivenInitials = str_sub(Given, 1, 2))

# Find CRAN packages by authors at ACEMS nodes
acems_places <- c("monash", "unimelb", "qut", "uq", "usyd", "uts", "unsw", "adelaide")
acemscran <- allcran %>%
  filter(stri_detect_regex(
    Maintainer,
    paste(paste0(acems_places, "\\.edu"), collapse = "|")
  ))

# Now select packages where maintainer is at ACEMS
acems_packages <- acemscran %>%
  inner_join(acems, by = c("GivenInitials", "Family")) %>%
  pull(Package)
# Add in other known packages
acems_packages <- c(
    acems_packages,
    "addb",
    "bayesImageS",
    "bfast",
    "BSL",
    "CEoptim",
    "colmozzie",
    "cricketdata",
    "curvHDR",
    "datapasta",
    "demography",
    "DescribeDisplay",
    "diffpriv",
    "dobson",
    "edrGraphicalTools",
    "eechidna",
    "emma",
    "expsmooth",
    "fable",
    "fds",
    "fma",
    "forecast",
    "fpp",
    "fpp2",
    "ftsa",
    "gammSlice",
    "glmmEP",
    "hdrcde",
    "hts",
    "LowRankQP",
    "MBSGS",
    "Mcomp",
    "MEFM",
    "mozzie",
    "naniar",
    "nullabor",
    "ozdata",
    "PIGE",
    "queuecomputer",
    "rainbow",
#    "rmarkdown",
    "robets",
    "rwalkr",
    "season",
    "serrsBayes",
    "sgPLS",
    "smoothAPC",
    "SSN",
    "staplr",
    "stR",
    "sugrrants",
    "taipan",
    "thief",
    "tourrGui",
    "tscompdata",
    "tsfeatures",
    "tsgeneration",
    "tsibble",
    "tsibbledata",
    "tsibblestats",
    "visdat"
  ) %>%
  unique() %>%
  sort()
acemscran <- allcran %>%
  filter(Package %in% acems_packages)
```

```{r downloads, echo = FALSE, message = FALSE, warning=FALSE}
# download data from CRAN
r_package_down <- cran_downloads(package = acems_packages,
                                 from = "2017-12-12",
                                 to = "2018-12-11")

# group by package and create count variable
r_packages <- r_package_down %>%
  group_by(package) %>%
  summarise(Total = sum(count)) %>%
  arrange(-Total) %>%
  rename(Package=package) %>%
  inner_join(acemscran, by="Package") %>%
  inner_join(acems, by=c("GivenInitials", "Family"))
```

One way ACEMS member translate their methodological research into practice is via R packages. R is an open-source language for statistical computing and data analysis. Research can be shared by putting R packages on the Comprehensive R Archive Network (CRAN), or on github, or some other public code repository. R packages allow anyone to implement cutting-edge research on their own data.

At the end of 2018, ACEMS members had contributed a total of `r NROW(acemscran)` packages on CRAN, plus many more on github. During 2018, the CRAN packages were downloaded a total of `r sprintf("%7.0f", sum(r_packages$Total))` times, and one (the `forecast` package by Rob Hyndman) was downloaded over 1 million times!  The table below shows all ACEMS packages that were downloaded more than 5000 times each during 2018.


```{r table}
r_packages %>%
  mutate(Maintainer = glue("{Given.x} {Family}")) %>%
  select(Package, Total, Maintainer, institution) %>%
  filter(Total > 5000) %>%
  rename(
    `Total downloads in 2018` = "Total",
    `ACEMS node` = "institution") %>%
  knitr::kable(booktabs=TRUE)
```

The following graph shows all packages authored by ACEMS members other than those downloaded more than 30,000 times during 2018.

```{r graph}
# generate ggplot using plotly
r_packages %>%
  filter(Total < 30000) %>%
  mutate(
    Package = reorder(Package, Total)) %>%
  ggplot(aes(x = Package, y=Total)) +
  geom_col() +
  coord_flip() +
  ggtitle("R Package Downloads for 2018")
```



